---
title: "Solving POMDPs with the appl R package"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Solving POMDPs with appl R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
## Don't run chunks in this vignette until it is ready.
knitr::opts_chunk$set(eval=FALSE)
```

```{r message=FALSE}
library(appl)
library(tidyverse) # for plotting
```

Parameters specifying our POMDP problem:

```{r}
states <- seq(0,1, length=50)
actions <- states
observations <- states
sigma_g <- 0.1
sigma_m <- 0.5
reward_fn <- function(x,h) pmin(x,h) # - .001*h
discount <- 0.95

r <- 1
K <- 0.75

f <- function(x, h){
  s <- pmax(x - h, 0)
  s * exp(r * (1 - s / K) )
}
```

A convenience function for generating transition, observation, and reward matrices given these parameters for the fisheries management problem:

```{r}
m <- fisheries_matrices(states, actions, observations, reward_fn, f, sigma_g, sigma_m, noise = "lognormal")
```

Long-running code to actually compute the solution.  Rather than run this, we'll load the solution from the log.

```{r eval=FALSE}

log_data <- data.frame(id = "vignette", model = "ricker", r = r, K = K, sigma_g = sigma_g, sigma_m = sigma_m)

system.time(
alpha <- sarsop(m$transition, m$observation, m$reward, discount, precision = .1, log_data = log_data, timeout = 20000)
)
```


Read a solution from the log:

```{r}
meta <- meta_from_log(data.frame(K = K, r = r, sigma_g = sigma_g, sigma_m = sigma_m, model = "ricker"))
alpha <- alphas_from_log(meta)[[1]]
```


Given the model matrices and `alpha` vectors and current belief state, compute & plot policy:

```{r}
df <- compute_policy(alpha, m$transition, m$observation, m$reward)
```

```{r}
ggplot(df, aes(states[state], states[state] - actions[policy])) + geom_line() + geom_point()
```


Simulate management under the POMDP policy:

```{r}
sim <- sim_pomdp(m$transition, m$observation, m$reward, discount, x0 = K, Tmax = 20, alpha = alpha)
```

Plot simulation

```{r}
sim$df %>%
  select(-value) %>%
  mutate(state = states[state], action = actions[action], obs = observations[obs]) %>%
  gather(variable, stock, -time) %>%
  ggplot(aes(time, stock, color = variable)) + geom_line()  + geom_point()
```

Plot belief evolution:

```{r}
Tmax <-length(out$state_posterior[,1])

out$state_posterior %>% data.frame(time = 1:Tmax) %>% 
  gather(state, probability, -time, factor_key =TRUE) %>% 
  mutate(state = as.numeric(state)) %>% 
  ggplot(aes(state, probability, group = time, alpha = time)) + geom_line()
```

